Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-13 paper/sections/proof-reconstruction.tex
[ paper ] minor changes.
[ paper ] resolve inference init.

\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ ¬ φ₂$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ ¬ φ₁ ∧ ¬ φ₂$}
\RightLabel{\tt ¬∧¬-to-¬∨}
\UnaryInfC{$Γ ⊢ ¬ (φ₁ ∨ φ₂)$}
\end{bprooftree}
\end{equation*}
where
\begin{equation*}
\mathcal{D} =
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}

(φ := ¬ (φ₁ ⇔ φ₂))
\begin{equation*}
\begin{bprooftree}

% -- right

\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂)) ∧ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁)) $}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁))$}
\RightLabel{\tt \thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₂ ⇒ \mysplit(φ₁)$}
\RightLabel{\tt weaken ¬ φ₂}
\UnaryInfC{$Γ , ¬ φ₂ ⊢ ¬ φ₂ ⇒ \mysplit(φ₁)$}

% -- right

% -- left
\AxiomC{}
\RightLabel{\tt assume (¬ φ₂)}
\UnaryInfC{$Γ , ¬ φ₂ ⊢ ¬ φ₂$}
\RightLabel{\tt ⇒-elim}

\BinaryInfC{$Γ , ¬ φ₂ ⊢ \mysplit(φ₁)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ , ¬ φ₂ ⊢ φ₁ $}
\RightLabel{\tt ⇒-intro}
\UnaryInfC{$Γ ⊢ ¬ φ₂ ⇒ φ₁$}
% -- left

% \RightLabel{\tt ∧-intro}
% \BinaryInfC{$Γ ⊢ (φ₁ ⇒ ¬ φ₂) ∧ (¬ φ₂ ⇒ φ₁)$}
% \RightLabel{\tt ⇒¬∧¬⇒-to-¬⇔}
% \UnaryInfC{$Γ ⊢ ¬ (φ₁ ⇔ φ₂)$}
\end{bprooftree}
\end{equation*}

@@

>>>>

\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ ¬ φ₂$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ ¬ φ₁ ∧ ¬ φ₂$}
\RightLabel{\tt ¬∧¬-to-¬∨}
\UnaryInfC{$Γ ⊢ ¬ (φ₁ ∨ φ₂)$}
\end{bprooftree}
\end{equation*}
where
\begin{equation*}
\mathcal{D} =
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
% (φ := ¬ (φ₁ ⇔ φ₂))
% \begin{equation*}
% \begin{bprooftree}

% % -- right

% \AxiomC{$Γ ⊢ \mysplit(φ, n)$}
% \RightLabel{By def.}
% \UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂)) ∧ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁)) $}
% \RightLabel{\tt ∧-proj₂}
% \UnaryInfC{$Γ ⊢ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁))$}
% \RightLabel{\tt \thmunshunt}
% \UnaryInfC{$Γ ⊢ ¬ φ₂ ⇒ \mysplit(φ₁)$}
% \RightLabel{\tt weaken ¬ φ₂}
% \UnaryInfC{$Γ , ¬ φ₂ ⊢ ¬ φ₂ ⇒ \mysplit(φ₁)$}

% % -- right

% % -- left
% \AxiomC{}
% \RightLabel{\tt assume (¬ φ₂)}
% \UnaryInfC{$Γ , ¬ φ₂ ⊢ ¬ φ₂$}
% \RightLabel{\tt ⇒-elim}

% \BinaryInfC{$Γ , ¬ φ₂ ⊢ \mysplit(φ₁)$}
% \RightLabel{\thmsplit}
% \UnaryInfC{$Γ , ¬ φ₂ ⊢ φ₁ $}
% \RightLabel{\tt ⇒-intro}
% \UnaryInfC{$Γ ⊢ ¬ φ₂ ⇒ φ₁$}
% % -- left

% % \RightLabel{\tt ∧-intro}
% % \BinaryInfC{$Γ ⊢ (φ₁ ⇒ ¬ φ₂) ∧ (¬ φ₂ ⇒ φ₁)$}
% % \RightLabel{\tt ⇒¬∧¬⇒-to-¬⇔}
% % \UnaryInfC{$Γ ⊢ ¬ (φ₁ ⇔ φ₂)$}
% \end{bprooftree}
% \end{equation*}

@@

\emph{Resolve.}
\newcommand{\thmresolve}{\name{thm-resolve}}
The \resolve rule is a generalization version for resolution
in propositional logic since the formulas are not shaped always as
the theorem expect (see Fig.~\ref{fig:metis-inferences}).
Since our approach is mere syntactic, in order to apply such a
theorem, we perform rearrangements inside the formulas to match with
the pattern expected by the resolution theorem. If this strategy
fails at some point, we could not check the rest of the proof.

Therefore, the best scenario to apply resolution occurs when the
literal $l$ and its negation appear in the exact place in the
formulas as the resolve inference shows.

\begin{figure}
\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ (φ₁ := l ∨ ψ₁)$}
\AxiomC{$Γ ⊢ (φ₂ := ¬ l ∨ ψ₂)$}
\RightLabel{\thmresolve}
\BinaryInfC{$Γ ⊢ \resolve(φ₁, φ₂, l, (ψ₁ ∧ ψ₂))$}
\UnaryInfC{Γ ⊢ ψ₁ ∨ ψ₂}
\end{bprooftree}
%}
\]
% \caption{Resolution in \Metis.}
\label{fig:thm-resolve-best-case}
\end{figure}

To rearrange the proposition in order to match with a specific
pattern of a formula, we define the following functions and provide
its respective theorems in \cite{AgdaMetis}.

\begin{itemize}
  \item[] \name{reorder-∨}
  \item[] \name{reorder-∧}
  \item[] \name{reorder-∧∨}
\end{itemize}

\begin{equation}
\resolve((φ₁, φ₂, l, ψ)

\begin{theorem}[\thmresolve] $Γ ⊢ φ₁$, $Γ ⊢ φ₂$,
then $Γ ⊢ \resolve(φ₁, φ₂, l, ψ)$ where $l$ is a literal,
and ψ is the expected result.
\end{theorem}

% \begin{proof}
% \begin{equation*}

% \end{equation*}
% \end{proof}

\emph{Clausification.}
